---
title: "R Forwards Chicago 2019 Package Workshop Evaluations"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}

library(ggplot2)
library(patchwork)
library(gganimate)
library(dplyr)
library(reshape)

knitr::opts_chunk$set(echo = TRUE)

pretest <- read.csv("R Forwards Package Workshop Learning Pre-Survey.csv")
posttest <- read.csv("R Forwards Package Workshop Learning Post-Survey.csv")

pretest$Which.of.the.following.things.describes.you. <- ifelse(pretest$Which.of.the.following.things.describes.you.
                                                               %in% c("Intermediate ", "Student;Young Professional","Young Professional;5 years experience", "Young Professional;Experienced Professional"), "Other", as.character(pretest$Which.of.the.following.things.describes.you.))

pretest$Which.of.the.following.things.describes.you. = factor(pretest$Which.of.the.following.things.describes.you.
                            ,levels= c("Student"
                                      , "Young Professional"
                                      , "Experienced Professional"
                                      , "Other"))


posttest$Which.of.the.following.things.describes.you. <- ifelse(posttest$Which.of.the.following.things.describes.you.
                                                               %in% c("Still intermediate", "Student;Young Professional","Young Professional;5 years experience", "Young Professional;Experienced Professional"), "Other", as.character(posttest$Which.of.the.following.things.describes.you.))

posttest$Which.of.the.following.things.describes.you. = factor(posttest$Which.of.the.following.things.describes.you.
                            ,levels= c("Student"
                                      , "Young Professional"
                                      , "Experienced Professional"
                                      , "Other"))
```

```{r, echo=FALSE}
# Rename columns
pretest_labs <- c("Timestamp","Can Find \n R Packages", "Can Use \n R Projects", "Has Github Acct"
                  , "Wrote \n Unit Test", "Regularly \n Work in R","Building \n R Functions"
                  , "Constructing \n R Packages"
                  , "Creating \n Unit Tests", "Using Git/Github", "Using R \n Projects"
                  , "Professional Level", "No Open \n Source Experience", "Chicago Local")

colnames(pretest) <- pretest_labs

```

```{r, echo=FALSE}
# Rename columns
posttest_labs <- c(pretest_labs, "Location \n Convenience", "Facilities", "Food"
                   , "Morning \n Teaching Skill", "Morning Content \n Expertise"
                   , "Afternoon \n Teaching Skill", "Afternoon Content \n Expertise"
                   , "Time for \n Practice", "Application \n Process", "Free Text Feedback")

colnames(posttest) <- posttest_labs

```

# Overview
This analysis represents the responses provided by attendees at the R Forwards Package Workshop
held in [Chicago on February 23, 2019](https://forwards.github.io/edu/chicago/).

About 30 women attended the workshop, and we have survey responses from 28 for the pre-survey 
and 29 for the post-survey.

## The Questions

We purposely designed this survey to have only questions that might be relevant
to improving the teaching of workshops like this in the future, and avoided asking
 for too much personal information about students.
 Because all responses on each survey are anonymous, we cannot connect the responses in
 the two surveys directly.

### Pre-test Items 
```{r, echo=FALSE}
colnames(pretest)
```

### Post-test Items
```{r, echo=FALSE}
colnames(posttest)
```




## Pre-test Descriptive Statistics {.tabset}

### Starting Characteristics
```{r, fig.width=8, fig.height=5, echo=FALSE}
p_tomelt <- pretest[, c(1:6)]
pretest_long <- melt(p_tomelt, id.vars = "Timestamp")

ggplot(pretest_long,aes(x=variable, fill=value))+
  theme_bw()+
  geom_bar()+
  labs(title = "Initial Skills Assessment", x="Skill")

```

```{r, fig.width=8, fig.height=5, echo=FALSE}
p_tomelt <- pretest[, c(1, 7:11)]
pretest_long <- melt(p_tomelt, id.vars = "Timestamp")
pretest_long$value = factor(pretest_long$value
                            ,levels= c("Very comfortable"
                                      , "Somewhat comfortable"
                                      , "Not very comfortable"
                                      , "Totally uncomfortable"
                                      , "I don't even know what that is"))

ggplot(pretest_long,aes(x=variable, fill=value))+
  theme_bw()+ 
  theme(legend.position="bottom")+guides(fill=guide_legend(nrow=2,byrow=TRUE))+
  geom_bar()+
  labs(title = "Comfort with Activities", x="Activity")

```

### Student Backgrounds
```{r, fig.width=8, fig.height=5, echo=FALSE}

p_tomelt <- pretest[, c(1, 12:14)]
pretest_long <- melt(p_tomelt, id.vars = "Timestamp")

plevel <- ggplot(pretest_long[pretest_long$variable == "Professional Level", ],aes(x=variable, fill=value))+
  theme_bw()+ 
  theme(legend.position="bottom")+guides(fill=guide_legend(nrow=2,byrow=TRUE))+
  geom_bar()+
  labs(title = "Background", x="Question")

tf <- ggplot(pretest_long[pretest_long$variable %in% c("Open Source Experience", "Chicago Local"), ],aes(x=variable, fill=value))+
  theme_bw()+ 
  theme(legend.position="bottom")+guides(fill=guide_legend(nrow=1,byrow=TRUE))+
  geom_bar()+
  labs(title = "Background", x="Question")

plevel + tf
```


### By Professional Level
Are there differences in these starting skills based on the individual's level of professional experience?

```{r, fig.width=10, fig.height=8, echo=FALSE}
p_tomelt <- pretest[, c(1, 12, 2:6)]
pretest_long <- melt(p_tomelt, id.vars = c("Timestamp","Professional Level"))

ggplot(pretest_long,aes(x=variable, fill=value))+
  facet_wrap(`Professional Level`~.)+
  theme_bw()+
  theme(legend.position="bottom")+
  geom_bar()+
  labs(title = "Initial Skills Assessment", x="Skill")

```

## Post-test Comparison {.tabset}

```{r, echo=FALSE}
pretest$testname <- "pre"
posttest$testname <- "post"

fullsample <- merge(pretest, posttest, by.x=colnames(pretest), by.y=colnames(pretest), all=T)

```


### Activities

```{r, echo=FALSE}

p_tomelt <- fullsample[, c(1,15, 7:11)]

fullsample_labs <- melt(p_tomelt, id.vars = c("Timestamp", "testname"))
fullsample_labs$value = factor(fullsample_labs$value
                            ,levels= c("Very comfortable"
                                      , "Somewhat comfortable"
                                      , "Not very comfortable"
                                      , "Totally uncomfortable"
                                      , "I don't even know what that is"))

fullsample_labs$testname = factor(fullsample_labs$testname
                            ,levels= c("pre"
                                      , "post"))

```

```{r, fig.width=10, fig.height=8, echo=FALSE}
ggplot(fullsample_labs,aes(x=variable, fill=value))+
  theme_bw()+ 
  theme(legend.position="bottom")+guides(fill=guide_legend(nrow=2,byrow=TRUE))+
  geom_bar(stat = 'count', position = 'stack') +
  labs(title = "Comfort with Activities: {closest_state}", x="Activity")+
  transition_states(
    testname,
    transition_length = 2,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('sine-in-out')

```

```{r, fig.width=10, fig.height=5, echo=FALSE}

ggplot(fullsample_labs,aes(x=variable, fill=value))+
  facet_grid(.~testname)+
  theme_bw()+ 
  theme(legend.position="bottom")+guides(fill=guide_legend(nrow=2,byrow=TRUE))+
  geom_bar(stat = 'count', position = 'stack') +
  labs(title = "Comfort with Activities", x="Activity")

```


### Activities by Professional Level


```{r, echo=FALSE}
p_tomelt <- fullsample[, c(1,15,12, 7:11)]

fullsample_labs <- melt(p_tomelt, id.vars = c("Timestamp", "testname", "Professional Level"))
fullsample_labs$value = factor(fullsample_labs$value
                            ,levels= c("Very comfortable"
                                      , "Somewhat comfortable"
                                      , "Not very comfortable"
                                      , "Totally uncomfortable"
                                      , "I don't even know what that is"))

fullsample_labs$testname = factor(fullsample_labs$testname
                            ,levels= c("pre"
                                      , "post"))


```

```{r, fig.width=10, fig.height=8, echo=FALSE}

ggplot(fullsample_labs,aes(x=variable, fill=value))+
  facet_wrap(`Professional Level`~.)+
  theme_bw()+ 
  theme(legend.position="bottom")+guides(fill=guide_legend(nrow=2,byrow=TRUE))+
  geom_bar(stat = 'count', position = 'stack') +
  labs(title = "Comfort with Activities: {closest_state}", x="Activity")+
  transition_states(
    testname,
    transition_length = 2,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('sine-in-out')
```


```{r, fig.width=10, fig.height=10, echo=FALSE}

ggplot(fullsample_labs,aes(x=variable, fill=value))+
  facet_wrap(`Professional Level`+testname~., ncol = 2)+
  theme_bw()+ 
  theme(legend.position="bottom")+guides(fill=guide_legend(nrow=2,byrow=TRUE))+
  geom_bar(stat = 'count', position = 'stack') +
  labs(title = "Comfort with Activities", x="Activity")

```

## Satisfaction with Program

```{r, fig.width=9, fig.height=7, echo=FALSE}

p_tomelt <- fullsample[, c(1, 15:24)]

fullsample_labs <- melt(p_tomelt, id.vars = c("Timestamp", "testname"))
fullsample_labs$value = factor(fullsample_labs$value
                            ,levels= c("Extremely Satisfied"
                                      , "Mostly Satisfied"
                                      , "It was just Ok"
                                      , "Mostly Unsatisfied"
                                      ))

fullsample_labs$testname = factor(fullsample_labs$testname
                            ,levels= c("pre"
                                      , "post"))


ggplot(fullsample_labs[fullsample_labs$testname == "post",],aes(x=variable, fill=value))+
  theme_bw()+ 
  theme(legend.position="bottom")+guides(fill=guide_legend(nrow=2,byrow=TRUE))+
  geom_bar(stat = 'count', position = 'stack') +
  labs(title = "Satisfaction", x="Item")

```

## Conclusions

Based on the responses shown, participants in the workshop clearly gained confidence 
in their abilities to work with R Packages and Unit Tests, among other things. 

We do know, from this data, that people would like more hands-on practice in the 
workshop, and there's some level of dissatisfaction with the convenience of our location 
- we were on the UChicago campus which is quite far from the city center. These are
lessons we will take and apply to future efforts.

Generally, this looks like a successful workshop and we are very happy with how 
everything went!
